Relative Stable Maps in Genus One via Central Alignments

### 2017-12-21 genus-one-relative-maps.tex
Some changes, made on master because I forgot. Sorry. Minor cosmetic changes anyway.

\title[Genus One Radial Maps]{Relative Stable Maps in Genus One via Radial Alignments}
\author{Luca Battistella and Navid Nabijou}
\begin{document}

@@

\section{Relative space equals closure of the nice locus}
Recall the definition of maps from centrally aligned curves via the cartesian diagram:
\bcd
\tVZc(X,\beta)\ar[d]\ar[r]\ar[dr,phantom,"\Box"] & \M{1}{n}{X}{\beta}\ar[d] \\
\MM_{1,n}^{\rm{ctr}}\ar[r] & \MM_{1,n}^{\dagger}
\ecd

\begin{dfn}
The centrally aligned relative space is defined within $\tVZc(Y,\beta)$ by the following:
\begin{itemize}
\item \emph{factorisation condition}: the map $f\colon C\to X$ factors through the associated contraction to a Smyth's singularity:
\bcd
\widetilde C\ar[r]\ar[d] & \overline C\ar[d,"\bar f"] \\
C\ar[r,"f"] & X
\ecd

\item \emph{Gathmann's relative condition}: for every connected component $Z$ of $f^{-1}(Y)$, either $Z$ is a point and, if it is marked $Z=x_i$, the multiplicity of $f$ at $x_i$ along $Y$ is at least $\alpha_i$; or $Z$ is a curve and $f^*\OO_X(Y)-\sum_{x_i\in Z}\alpha_ix_i$ is effective. Notice that $Z$ is at most genus $1$ and every line bundle of positive degree on a Gorenstein irreducible genus $1$ curve is effective, hence this condition can be rephrased as: the numerical condition
\[f_*[Z]\cdot Y+\sum_{j=1}^r m^{(j)}\geq \sum_{x_i\in Z}\alpha_i,\]
and, in case the numerical equality holds, the following equality of line bundles:
\[f_{\lvert Z}^*\OO_X(Y)=\OO_Z\left(\sum_{x_i\in Z}\alpha_ix_i - \sum_{j=1}^r m^{(j)}y_j\right)\in\Pic(Z).\]

\item \emph{novel condition}: there are no two vertices of $\rm{MinPos}$ closest to the circuit (i.e. $v_1\neq v_2\in \rm{MinPos}$ with $\delta=\lambda(v_1)=\lambda(v_2)$) unless the corresponding $m^{(j_1)}$ and $m^{(j_2)}$ are equal.
\end{itemize}
\end{dfn}

Here is one example where we show by a dimensional computation that the novel condition must be included if we hope to determine the closure of the nice locus.
\begin{example}
Consider $\M{1}{(3)}{\PP^1|H}{3}$. The virtual dimension is $7-3=4$. Here is a parametrisation of the nice locus: choose an element $(E,p)$ of $\oM_{1,1}\setminus\partial\oM_{1,1}$ (which has dimension $1$), let then $s_0\colon\OO_E\hookrightarrow\OO_E(3p)$ and $s_1$ be any section of $\OO_E(3p)$ not vanishing at $p$ (notice that $h^0(E,\OO_E(3p))=3$). Consider now the following weighted graph for a map in the boundary:
\begin{center}
\begin{tikzpicture}
\draw[color=brown] (0,0) node[left]{$E$} -- (4,0);
\draw[fill=black] (1,0) circle[radius=2pt] node[above]{$x_1$};
\draw[fill=black] (2,0) circle[radius=2pt] node[below right]{$y_1$};
\draw[fill=black] (3,0) circle[radius=2pt] node[below right]{$y_2$};
\draw (2,1.5) node[left]{$R_1$} (3,1.5) node[left]{$R_2$};
\draw (2,-1) -- (2,2) node[above]{$2$} (3,-1) -- (3,2) node[above]{$1$};
\draw[->] (5,0.5) -- (8,0.5);
\draw (9,-1) -- (9,2) node[above]{$\PP^1$};
\draw[fill=black] (9,0) circle[radius=2pt] node[right]{$H$};
\end{tikzpicture}
\end{center}
where the brown line represents a contracted genus $1$ curve. Now $(E,x_1,y_1,y_2)$ is a point of $\oM_{1,3}$ subject to the divisorial condition $3x_1-2y_1-y_2=0\in A_0(E)$; furthermore we have to choose the second ramification value of the $2\colon 1$ map from $R_1$ to $\PP^1$. This already makes up for a $3$-dimensional moduli space of degenerate relative maps corresponding to such a graph. The minimal log structure for this curve has a chart from $\-mathbb N^2$, with generators $e_1$ and $e_2$ corresponding to the smoothing parameters of the two nodes. If we allowed $e_1$ and $e_2$ to be identified in the characteristic sheaf, then we would get an extra $\Gm$ of choices for the log structure, so in total a $4$-dimensional moduli space. There is no chance this could live in the closure of the nice locus.
\end{example}

>>>>

\title[Genus One Radial Maps]{Relative Stable Maps in Genus One via Central Alignments}
\author{Luca Battistella and Navid Nabijou}
\begin{document}

@@

\section{Relative space equals closure of the nice locus}
Recall that the moduli space of maps from centrally aligned curves is defined via the following Cartesian diagram:
\bcd
\tVZc(X,\beta)\ar[d]\ar[r]\ar[dr,phantom,"\Box"] & \M{1}{n}{X}{\beta}\ar[d] \\
\MM_{1,n}^{\rm{ctr}}\ar[r] & \MM_{1,n}^{\dagger}
\ecd

\begin{dfn}
The \emph{centrally aligned relative space} is defined to be the substack of $\tVZc(Y,\beta)$ satisfying the following conditions:
\begin{enumerate}
\item \emph{Factorisation condition.} The map $f\colon C\to X$ factors through the associated contraction to a Smyth singularity:
\bcd
\widetilde C\ar[r]\ar[d] & \overline C\ar[d,"\bar f"] \\
C\ar[r,"f"] & X
\ecd

\item \emph{Gathmann's relative condition.} Let $Z$ be a connected component of $f^{-1}(Y)$. If $Z$ is a marked point $x_i$ then the multiplicity of $f$ at $x_i$ along $Y$ is at least $\alpha_i$. On the other hand if $Z\subseteq C$ is a curve then
\begin{equation*} f^*\OO_X(Y)|_Z -\sum_{x_i\in Z}\alpha_ix_i\end{equation*}
is an effective class on $Z$. Notice that since $Z$ is at most genus $1$ and every line bundle of positive degree on a Gorenstein irreducible genus $1$ curve is effective, this condition can be rephrased as a numerical condition
\begin{equation*} f_*[Z]\cdot Y+\sum_{j=1}^r m^{(j)}\geq \sum_{x_i\in Z}\alpha_i \end{equation*}
together with the following equality of line bundles in $\Pic(Z)$:
\begin{equation*} (f|_{Z})^*\OO_X(Y)=\OO_Z\left(\sum_{x_i\in Z}\alpha_ix_i - \sum_{j=1}^r m^{(j)}y_j\right)\end{equation*}

\item \emph{Novel condition.} There are no two vertices of $\rm{MinPos}$ closest to the circuit (i.e. $v_1, v_2\in \rm{MinPos}$  with $\delta=\lambda(v_1)=\lambda(v_2)$) unless the corresponding $m^{(j_1)}$ and $m^{(j_2)}$ are equal.
\end{enumerate}
\end{dfn}

Remember that we are trying to impose conditions which determine the closure of the nice locus inside the moduli space of centrally aligned maps. Here is one example where we show by a dimensional computation that the novel condition must be included if we hope to determine the closure of the nice locus.
\begin{example}
Consider $\M{1}{(3)}{\PP^1|H}{3}$. The virtual dimension is $7-3=4$. Here is a parametrisation of the nice locus: choose an element $(E,p)$ of $\oM_{1,1}\setminus\partial\oM_{1,1}$ (which has dimension $1$), let then $s_0\colon\OO_E\hookrightarrow\OO_E(3p)$ and $s_1$ be any section of $\OO_E(3p)$ not vanishing at $p$ (notice that $h^0(E,\OO_E(3p))=3$). Consider now the following weighted graph for a map in the boundary:
\begin{center}
\begin{tikzpicture}
\draw[color=brown] (0,0) node[left]{$E$} -- (4,0);
\draw[fill=black] (1,0) circle[radius=2pt] node[above]{$x_1$};
\draw[fill=black] (2,0) circle[radius=2pt] node[below right]{$y_1$};
\draw[fill=black] (3,0) circle[radius=2pt] node[below right]{$y_2$};
\draw (2,1.5) node[left]{$R_1$} (3,1.5) node[left]{$R_2$};
\draw (2,-1) -- (2,2) node[above]{$2$} (3,-1) -- (3,2) node[above]{$1$};
\draw[->] (5,0.5) -- (8,0.5);
\draw (9,-1) -- (9,2) node[above]{$\PP^1$};
\draw[fill=black] (9,0) circle[radius=2pt] node[right]{$H$};
\end{tikzpicture}
\end{center}
where the brown line represents a contracted genus $1$ curve. Now $(E,x_1,y_1,y_2)$ is a point of $\oM_{1,3}$ subject to the divisorial condition $3x_1-2y_1-y_2=0\in A_0(E)$; furthermore we have to choose the second ramification value of the $2\colon 1$ map from $R_1$ to $\PP^1$. This already makes up for a $3$-dimensional moduli space of degenerate relative maps corresponding to such a graph. The minimal log structure for this curve has a chart from $\-mathbb N^2$, with generators $e_1$ and $e_2$ corresponding to the smoothing parameters of the two nodes. If we allowed $e_1$ and $e_2$ to be identified in the characteristic sheaf, then we would get an extra $\Gm$ of choices for the log structure, so in total a $4$-dimensional moduli space. There is no chance this could live in the closure of the nice locus.
\end{example}
